---
title: "R-code for 'Survival of common buzzards: Effects of condition, parasites, and climate on individual-specific rates'"
author: "Meinolf Ottensmann, Anja Wiegmann, Tony Rinaud, Oliver Krüger, Christina Strube, Jamie Winternitz and Nayden Chakarov"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
library(knitr)
knitr::opts_chunk$set(
  fig.width = 8, fig.height = 8,highlight = TRUE, comment = "#>",
  strip.white = TRUE,collapse = TRUE, fig.align = "center", cache = TRUE,
  warning = F, message = F)
```


This document provides the `R code` for a supplementary analysis of the Capture-Mark-Resighting data using a simpler `Burnham`model. See the RMarkdown file `01_R_code` for more background on the main analyses presented in our paper. 

```{r, message=FALSE, warning=FALSE}
if ("EnvStats" %in% rownames(installed.packages())) {
  library("EnvStats")
} else {
  install.packages("EnvStats")
}
if ("magrittr" %in% rownames(installed.packages())) {
  library("magrittr")
} else {
  install.packages(magrittr)
}
if ("patchwork" %in% rownames(installed.packages())) {
  library("patchwork")
} else {
  install.packages(patchwork)
}
if ("RMark" %in% rownames(installed.packages())) {
  library("RMark")
} else {
  install.packages(RMark)
}
if ("tidyverse" %in% rownames(installed.packages())) {
  library("tidyverse")
} else {
  install.packages(tidyverse)
}
if ("egg" %in% rownames(installed.packages())) {
  library("egg")
} else {
  install.packages(egg)
}
```

Furthermore, a number of custom scripts are required. All of these are available within the file `R/custom_functions.R`. The function `source()` enables to use them.

```{r}
source("R/custom_functions.R")
```

Individual-specific data are available as `ringing_data` (i.e., details on ringing dates and individual-specific covariates) and `resighting_data` (i.e., details on live encounters).

```{r, message=FALSE}
ringing_data <- as.data.frame(read_csv("data/ringing_data.csv")) 
resighting_data <- as.data.frame(read_csv("data/resighting_data.csv")) %>% 
  filter(Ring %in% ringing_data$Ring)
```

Additionally, we use North Atlantic oscillation (`NAO`) indices that are based on the principal component approach [\^8], retrieved from <https://climatedataguide.ucar.edu/climate-data/hurrell-north-atlantic-oscillation-nao-index-pc-based> (last changed 2021-08-28). As recording intervals cross years (May - April), NAO indices for the subsequent year need to be assigned: E.g., 2007 and DJFM 2008.

Note, for each recording interval, DJFM values of the subsequent year are used (e.g., cohort 2007 = NAO in 2008). This is taken care of by defining years

```{r}
nao <- read.table("data/nao.txt", header = T) %>% 
  mutate(time = as.factor(year - 1))
```

Add climate data from local weather stations

```{r}
climate.data <- readRDS("data/climate.data.RDS")
climate.data$season$time <- as.factor(climate.data$season$time)
nao <- left_join(nao, climate.data$season, by = "time")
```

## Modelling of survival probabilities
### Burnham model (live-dead encounters)

We model survival probabilities using combined dead and live encounter events using the `Burnham` model implemented in software `MARK`. This model does not allow missing covariates and hence we need to subset the dataset first (i.e., removing the cohort of 2008 where no data on *Leucocytozoon* infection status is available):

```{r}
## Subset to wing-tagged birds with complete covariates
## ----------------------------------------------------
ringing_data <- filter(
  ringing_data,
  # Subset to wing-tagged only
  Tag ==  1, 
  # Remove birds that died in the nest
  !CauseDeath %in% "Dead as chick",
  # Remove two birds predated as nestling
  !Ring %in% c("3408543", "3417787"),
  # known Sex
  !is.na(Sex),
  # No missing covariate 
  !is.na(Lprev),
  # Keep birds with assigned morph
  Morph %in% 2:4) %>% 
  #mutate(Rank = case_when(Rank == "1" ~ "A", Rank != "1" ~ "B")) %>% 
  mutate(Lbinom = as.factor(ifelse(Lprev == 0, "Uninfected", "Infected"))) %>% 
  ## scale continuous variables
  mutate(Hatch = scale(Hatch)) %>% 
  mutate(Age_days = scale(Age_days)) %>% 
  mutate(BCI = scale(BCI))
```

At this stage, we retain a total of `r nrow(ringing_data)` individuals, corresponding to a mean cohort size of `r mean(table(ringing_data$Year))` [Range: `r min(table(ringing_data$Year))` - `r max(table(ringing_data$Year))`].

Based on ringing data and re-sightings of individuals, we assemble combined live-dead encounter histories for all individuals using the custom function `create.ld`. This functions generates a vector of the form `LDLDLD`, where `L` codes captures (here, only the marking occasion) and `D` codes both dead recoveries (D = 1) and live re-sightings (D = 2) between sampling intervals. Together, each `LD` pair codes for one interval of the study period. Next, we define the model structure for the `Burnham` model and code the required design data.

```{r, eval=FALSE}
## NOT run, output is saved
## Code capture histories (ch) and reformat data
## ------------------------------------------------------
df.burnham <- data.frame(
  ringing_data,
  ch =  create.ld(
    start = 2007,
    end = 2020, 
    ringing_data = ringing_data,
    resights = resighting_data)) %>% 
  mutate(ch = as.character(ch)) %>% 
  mutate(Morph = factor(Morph, labels = c("D","I","L"))) %>% 
  mutate(Sex = as.factor(Sex)) %>%
  subset(., select = -c(Ring, Tag, Dead, DateDeath, CauseDeath)) 

## process the data for RMark
## ------------------------------------------------------
dp.burnham  <- process.data(data = df.burnham, model = "Burnham", begin.time = 2007,
                            groups = c("Sex", "Morph", "Lbinom"))

## create design data
## ------------------------------------------------------
ddl.burnham <- make.design.data(dp.burnham) 

## Bin age-classes
## Survival S:
## ------------------------------------------------------ 
ddl.burnham <- add.design.data(
  data = dp.burnham, ddl = ddl.burnham, parameter="S", type = "age",
  name = "age", bins = c(0, 1, 2, 13),
  right = FALSE, replace = TRUE)

## Add dummy variable to restrict some factors to affect only first
## certain age classes
## ---------------------------------------------------------------
## First year
ddl.burnham$S$y1 <- 0
ddl.burnham$S$y1[ddl.burnham$S$age == "[0,1)"] <- 1

## Recapture probability p:
## ------------------------------------------------------
ddl.burnham <- add.design.data(
  data = dp.burnham, ddl = ddl.burnham, parameter = "p", type = "age",
  name = "age", bins = c(0, 1, 2, 13),
  right = FALSE, replace = TRUE)
ddl.burnham$p <- ddl.burnham$p %>% 
  mutate(ageclass2 = case_when(
    age == "[1,2)" ~ "CY2",
    age != "[1,2)" ~ "CY3plus"
  ))

## Recovery probability r:
## ------------------------------------------------------
ddl.burnham <- add.design.data(
  data = dp.burnham, ddl = ddl.burnham, parameter = "r", type = "age",
  name = "age", bins = c(0, 1, 2, 13),
  right = FALSE, replace = TRUE)
ddl.burnham$r <- ddl.burnham$r %>% 
  mutate(ageclass2 = case_when(
    age == "[0,1)" ~ "CY1",
    age != "[0,1)" ~ "CY2plus"
  ))

## collect variables which are year but not individual specific
## ------------------------------------------------------
covars <- left_join(data.frame(time = ddl.burnham$S$time), nao, "time")

## extend design data with NAO indices
## ------------------------------------------------------
ddl.burnham$S$NAO <- covars$DJFM

## Save files for later use 
## -----------------------------------------------------------------------------
saveRDS(df.burnham, file = "data/df.burnham.RDS", compress = "bzip2")
saveRDS(dp.burnham, file = "data/dp.burnham.RDS", compress = "bzip2")
saveRDS(ddl.burnham, file = "data/ddl.burnham.RDS", compress = "bzip2")
```

Load data previously aggregated using the above code

```{r}
df.burnham <- readRDS("data/df.burnham.RDS")
dp.burnham <- readRDS("data/dp.burnham.RDS")
ddl.burnham <- readRDS("data/ddl.burnham.RDS")
```

#### Estimating c-hat based on bootstrapping

```{r, eval = F, message=FALSE, warning=FALSE}
## Run general model to estimate parameters for simulations
## NOT run, output is saved for computational reasons
## --------------------------------------------------------
global.model <- mark(
  dp.burnham,  ddl.burnham, silent = FALSE, output = FALSE, invisible = TRUE,
  model.parameters = list(
    S = list(formula = ~ age + time + Sex, 
             fixed = list(cohort = 2008, value = 1)), 
    p = list(formula = ~ time, 
             fixed = list(cohort = 2008, value = 0)),
    F = list(formula = ~ 1, fixed = 1),
    r = list(formula = ~ ageclass2, 
             fixed = list(cohort = 2008, value = 0))),
  model = 'Burnham')
cleanup(ask = FALSE)

## Extract estimated parameters
## --------------------------------------------------------
params <- lapply(row.names(global.model$results$real), function(x) {
  stringr::str_split(x, pattern = " ")[[1]]
}) %>% 
  do.call("rbind",.) %>% 
  as.data.frame() %>% 
  set_colnames(c("Parameter", "Group", "Cohort", "age", "time")) %>% 
  mutate(time = substr(time, 2, nchar(time)) %>% as.numeric()) %>% 
  mutate(age = substr(age, 2, nchar(age)) %>% as.numeric()) %>% 
  cbind(., global.model$results$real) 

## Split by parameter (S, r, p & F) and group (Female, Male)
## --------------------------------------------------------

## Survival:
## --------------------------------------------------------
S.f <- filter(params, Group == "gFemaleDInfected", Parameter == "S") 
S.m <- filter(params, Group == "gMaleDInfected", Parameter == "S")

## Recovery probability
## --------------------------------------------------------
r <- filter(params, Parameter == "r")

## Re-sighting probability.
## --------------------------------------------------------
p <- filter(params, Parameter == "p")

## Site-fidelity is fixed to 1!
## --------------------------------------------------------
F <- 1

## get group table to assign parameters correctly in the bs model
## --------------------------------------------------------
(group.table <- filter(dp.burnham$data[,c("group", "Sex", "Morph", "Lbinom")]) %>% 
   unique.data.frame() %>% 
   .[with(., order(Sex, Morph, Lbinom)),])


## conduct simulations using parametric bs for both models
## -----------------------------------------------------------------------------
sim.out <- sims.burnham(
  S.fem = S.f, S.mal = S.m, p.fem = p, p.mal = p,
  F.fem = F, F.mal = F, r.fem = r, r.mal = r,
  marked =  Marked(
    data = dp.burnham$data, n.occasions = nchar(dp.burnham$data$ch[1])/2,
    groups = group.table$group, group.col = "group"),
  reps = 999, 
  model.parameters = list(
    S = list(formula = ~ time + age + Sex, 
             fixed = list(cohort = 2008, value = 1)), 
    p = list(formula = ~ time, 
             fixed = list(cohort = 2008, value = 0)),
    F = list(formula = ~ 1, fixed = 1),
    r = list(formula = ~ ageclass2, 
             fixed = list(cohort = 2008, value = 0))))

save(sim.out, file = "data/sim.out.burnham.RData")
c.hat <- list(median = global.model$results$deviance/sim.out$deviance.median,
              lcl = global.model$results$deviance/sim.out$deviance.025,
              ucl = global.model$results$deviance/sim.out$deviance.975)
save(c.hat, file = "data/c.hat.burnham.RData")
cleanup(ask = FALSE)
```

Inspect results of the bootstrapping exercise:

```{r}
load("data/sim.out.burnham.RData")
load("data/c.hat.burnham.RData")

ggplot(data.frame(x = (c.hat$median*sim.out$deviance.median)/sim.out$deviance.raw), aes(x)) +
  geom_histogram(bins = 30, col = "black", fill = "white") +
  theme_article() +
  xlab(TeX("Estimated $\\hat{C}")) +
  ylab("Frequency") +
  annotate("text", x = -Inf, y = Inf, hjust = -0.2, vjust = 1.5, 
           label = TeX(paste0("$\\hat{C} = ", round(c.hat$median,2)),
                       output = 'character'),parse = T) +
  annotate("text", x = -Inf, y = Inf, hjust = -0.12, vjust = 4, 
           label = paste("CI:", round(c.hat$ucl,2), "-", round(c.hat$lcl,2)))
```

#### Candidate models

```{r, eval=FALSE}
do_analysis.burnham = function(dp.burnham, ddl.burnham) {
  ## Models for Survival S
  ## ---------------------------------------------------------------------------
  
  ## ---------------------------------------------------------------------------
  S.01 = list(formula = ~ age, fixed = list(cohort = 2008, value = 1))
  S.02 = list(formula = ~ age + time, fixed = list(cohort = 2008, value = 1))
  S.03 = list(formula = ~ age + NAO, fixed = list(cohort = 2008, value = 1))
  S.04 = list(formula = ~ age * NAO, fixed = list(cohort = 2008, value = 1))
  S.05 = list(formula = ~ age + NAO + time, fixed = list(cohort = 2008, value = 1))
  S.06 = list(formula = ~ age * NAO + time, fixed = list(cohort = 2008, value = 1))
  
  ## Sex
  ## ---------------------------------------------------------------------------
  S.07 = list(formula = ~ age + Sex, fixed = list(cohort = 2008, value = 1))
  S.08 = list(formula = ~ age + time + Sex, fixed = list(cohort = 2008, value = 1))
  S.09 = list(formula = ~ age + NAO + Sex, fixed = list(cohort = 2008, value = 1))
  S.10 = list(formula = ~ age * NAO + Sex, fixed = list(cohort = 2008, value = 1))
  S.11 = list(formula = ~ age + NAO + time + Sex, fixed = list(cohort = 2008, value = 1))
  S.12 = list(formula = ~ age * NAO + time + Sex, fixed = list(cohort = 2008, value = 1))
  
  ## Morph
  ## ---------------------------------------------------------------------------
  S.13 = list(formula = ~ age + Morph, fixed = list(cohort = 2008, value = 1))
  S.14 = list(formula = ~ age + time + Morph, fixed = list(cohort = 2008, value = 1))
  #S.17 = list(formula = ~ age * time + Morph, fixed = list(cohort = 2008, value = 1))
  S.15 = list(formula = ~ age + NAO + Morph, fixed = list(cohort = 2008, value = 1))
  S.16 = list(formula = ~ age * NAO + Morph, fixed = list(cohort = 2008, value = 1))
  S.17 = list(formula = ~ age + NAO + time + Morph, fixed = list(cohort = 2008, value = 1))
  S.18 = list(formula = ~ age * NAO + time + Morph, fixed = list(cohort = 2008, value = 1))
  ## Leucocytozoon
  ## ---------------------------------------------------------------------------
  S.19 = list(formula = ~ age + Lbinom, fixed = list(cohort = 2008, value = 1))
  S.20 = list(formula = ~ age + time + Lbinom, fixed = list(cohort = 2008, value = 1))
  S.21 = list(formula = ~ age + NAO + Lbinom, fixed = list(cohort = 2008, value = 1))
  S.22 = list(formula = ~ age * NAO + Lbinom, fixed = list(cohort = 2008, value = 1))
  S.23 = list(formula = ~ age + NAO + time + Lbinom, fixed = list(cohort = 2008, value = 1))
  S.24 = list(formula = ~ age * NAO + time + Lbinom, fixed = list(cohort = 2008, value = 1))
  
  S.25 = list(formula = ~ age + y1:Lbinom, fixed = list(cohort = 2008, value = 1))
  S.26 = list(formula = ~ age + time + y1:Lbinom, fixed = list(cohort = 2008, value = 1))
  S.27 = list(formula = ~ age + NAO + y1:Lbinom, fixed = list(cohort = 2008, value = 1))
  S.28 = list(formula = ~ age * NAO + y1:Lbinom, fixed = list(cohort = 2008, value = 1))
  S.29 = list(formula = ~ age + NAO + time + y1:Lbinom, fixed = list(cohort = 2008, value = 1))
  S.30 = list(formula = ~ age * NAO + time + y1:Lbinom, fixed = list(cohort = 2008, value = 1))
  
  S.31 = list(formula = ~ age + y1:Age_days:Lbinom, fixed = list(cohort = 2008, value = 1))
  S.32 = list(formula = ~ age + time + y1:Age_days:Lbinom, fixed = list(cohort = 2008, value = 1))
  #S.38 = list(formula = ~ age * time + y1:Age_days:Lbinom, fixed = list(cohort = 2008, value = 1))
  S.33 = list(formula = ~ age + NAO + y1:Age_days:Lbinom, fixed = list(cohort = 2008, value = 1))
  S.34 = list(formula = ~ age * NAO + y1:Age_days:Lbinom, fixed = list(cohort = 2008, value = 1))
  S.35 = list(formula = ~ age + NAO + time + y1:Age_days:Lbinom, fixed = list(cohort = 2008, value = 1))
  S.36 = list(formula = ~ age * NAO + time + y1:Age_days:Lbinom, fixed = list(cohort = 2008, value = 1))
  
  ## Body condition
  ## ---------------------------------------------------------------------------
  S.37 = list(formula = ~ age + BCI, fixed = list(cohort = 2008, value = 1))
  S.38 = list(formula = ~ age * BCI, fixed = list(cohort = 2008, value = 1))
  S.39 = list(formula = ~ age + time + BCI, fixed = list(cohort = 2008, value = 1))
  S.40 = list(formula = ~ age * BCI + time, fixed = list(cohort = 2008, value = 1))
  S.41 = list(formula = ~ age + NAO + BCI, fixed = list(cohort = 2008, value = 1))
  S.42 = list(formula = ~ age * BCI + NAO, fixed = list(cohort = 2008, value = 1))
  S.43 = list(formula = ~ age + NAO + time + BCI, fixed = list(cohort = 2008, value = 1))
  S.44 = list(formula = ~ age * BCI + time + NAO, fixed = list(cohort = 2008, value = 1))
  
  S.45 = list(formula = ~ age + y1:BCI, fixed = list(cohort = 2008, value = 1))
  S.46 = list(formula = ~ age + time + y1:BCI, fixed = list(cohort = 2008, value = 1))
  S.47 = list(formula = ~ age + NAO + y1:BCI, fixed = list(cohort = 2008, value = 1))
  S.48 = list(formula = ~ age + NAO + time + y1:BCI, fixed = list(cohort = 2008, value = 1))
  
  S.49 = list(formula = ~ age * NAO + y1:BCI, fixed = list(cohort = 2008, value = 1))
  S.50 = list(formula = ~ age * NAO + time + y1:BCI, fixed = list(cohort = 2008, value = 1))
  
  ## Rank
  ## ---------------------------------------------------------------------------
  S.51 = list(formula = ~ age + y1:Rank, fixed = list(cohort = 2008, value = 1))
  S.52 = list(formula = ~ age + time + y1:Rank, fixed = list(cohort = 2008, value = 1))
  S.53 = list(formula = ~ age + NAO + y1:Rank, fixed = list(cohort = 2008, value = 1))
  S.54 = list(formula = ~ age * NAO + y1:Rank, fixed = list(cohort = 2008, value = 1))
  S.55 = list(formula = ~ age + NAO + time + y1:Rank, fixed = list(cohort = 2008, value = 1))
  S.56 = list(formula = ~ age * NAO + time + y1:Rank, fixed = list(cohort = 2008, value = 1))
  
  ## Hatching date
  ## ---------------------------------------------------------------------------
  S.57 = list(formula = ~ age + y1:Hatch, fixed = list(cohort = 2008, value = 1))
  S.58 = list(formula = ~ age + time + y1:Hatch, fixed = list(cohort = 2008, value = 1))
  S.59 = list(formula = ~ age + NAO + y1:Hatch, fixed = list(cohort = 2008, value = 1))
  S.60 = list(formula = ~ age * NAO + y1:Hatch, fixed = list(cohort = 2008, value = 1))
  S.61 = list(formula = ~ age + NAO + time + y1:Hatch, fixed = list(cohort = 2008, value = 1))
  S.62 = list(formula = ~ age * NAO + time + y1:Hatch, fixed = list(cohort = 2008, value = 1))
  
  ## Age at sampling date
  ## ---------------------------------------------------------------------------
  S.63 = list(formula = ~ age + y1:Age_days, fixed = list(cohort = 2008, value = 1))
  S.64 = list(formula = ~ age + time + y1:Age_days, fixed = list(cohort = 2008, value = 1))
  S.65 = list(formula = ~ age + NAO + y1:Age_days, fixed = list(cohort = 2008, value = 1))
  S.66 = list(formula = ~ age * NAO + y1:Age_days, fixed = list(cohort = 2008, value = 1))
  S.67 = list(formula = ~ age + NAO + time + y1:Age_days, fixed = list(cohort = 2008, value = 1))
  S.68 = list(formula = ~ age * NAO + time + y1:Age_days, fixed = list(cohort = 2008, value = 1))
  
  ## Models for re-sighting probability p
  ## p.time.age: Not supported
  ## ---------------------------------------------------------------------------
  p.dot = list(formula = ~ 1, fixed = list(cohort = 2008, value = 0))
  p.time = list(formula = ~ time, fixed = list(cohort = 2008, value = 0))

  ## Models for Site fidelity F
  ## ---------------------------------------------------------------------------
  F.fixed = list(formula = ~ 1, fixed = 1)

    ## Models for dead recovery chances r
  ## ---------------------------------------------------------------------------
  r.dot = list(formula = ~ 1, fixed = list(cohort = 2008, value = 0))
  r.age2 = list(formula = ~ ageclass2, fixed = list(cohort = 2008, value = 0))

  ## Compile model list based on parameter combinations
  ## ---------------------------------------------------------------------------
  cml = create.model.list(model = "Burnham")
  
  ## Run models
  ## ---------------------------------------------------------------------------
  results.burnham = mark.wrapper.parallel(
    model.list = cml, data = dp.burnham, ddl = ddl.burnham,
    output = FALSE, silent = FALSE, cpus = 2, parallel = TRUE)
  return(results.burnham)
}

## Conduct analysis
## ---------------------------------------------------------------------------
models.burnham <- do_analysis.burnham(dp.burnham =  dp.burnham, ddl.burnham = ddl.burnham)
#saveRDS(models.burnham, file = "data/models.burnham.RDS", compress = 'bzip2')

## adjust for over dispersion using c-hat
## ---------------------------------------------------------------------------
load("data/c.hat.burnham.RData")
models.burnham.adj <-  adjust.chat(model.list = models.burnham, chat = c.hat$median) 
saveRDS(models.burnham.adj, file = "data/models.burnham.adj.RDS", compress = 'bzip2')

## Select supported models
## ---------------------------------------------------------------------------
Large.AICc <- 
  rownames(models.burnham.adj$model.table)[which(models.burnham.adj$model.table[["DeltaQAICc"]] > 7)] %>% 
  as.numeric()

supported.models.burnham <- remove.mark(
  marklist = models.burnham.adj,
  model.numbers = Large.AICc)
saveRDS(supported.models.burnham, file = "data/supported.models.burnham.RDS")
cleanup(ask = FALSE)
unlink(list.files(pattern = "mark", ignore.case = F))
```

#### Results

```{r}
## all models
#models.burnham <- readRDS("data/models.burnham.RDS")
## models with adjusted deviance
models.burnham.adj <- readRDS("data/models.burnham.adj.RDS")
## models within Delta_QAICc <= 10 of the top ranking
supported.models.burnham <- readRDS("data/supported.models.burnham.RDS")
```

```{r, eval=FALSE, echo=FALSE}
## create model table
models <- data.frame(
  model.number = 1:(length(models.burnham.adj) - 1),
  model = sapply(models.burnham.adj[1:(length(models.burnham.adj) - 1)], function(x) x[["model.name"]]))

#View(supported.models.burnham$model.table)
```


##### Model averaging

Models including age, time, NAO and an additional covariate (BCI or *Leucocytozoon* prevalence) are not fully estimatable due to boundary estimates! For survival estimates this is the case for the following two models that are thus not considered in model averaging.


```{r}
S.mod.avg <- model.average(supported.models.burnham,
                           parameter = "S", drop = T, vcv = T)[["estimates"]] %>%
  ## estimates 2-13 are identical, but age == 3 needed for 2010!
  filter(fixed != "Fixed", age %in% 0:3) %>% 
  subset(., select = c(estimate, se, lcl, ucl, age, Sex, Morph, Lbinom)) %>% 
  unique.data.frame()

## Model average recovery estimates (r)
## -----------------------------------------------------------------------------
r.mod.avg <- model.average(supported.models.burnham,
                           parameter = "r", drop = T, vcv = T)[["estimates"]] %>% 
  filter(fixed != "Fixed") %>% 
  subset(., select = (c(estimate, se, lcl, ucl, age, Sex))) %>% 
  unique.data.frame()

## Model average resighting estimates (p)
## -----------------------------------------------------------------------------
p.mod.avg <- model.average(supported.models.burnham,
                           parameter = "p", drop = T, vcv = T)[["estimates"]] %>% 
  filter(fixed != "Fixed") %>% 
  subset(., select = (c(estimate, se, lcl, ucl, time, age))) %>% 
  unique.data.frame()
```


##### Model selection

```{r, eval=FALSE}
## Format model table
## -----------------------------------------------------------------------------
model.table <- 
  models.burnham.adj$model.table[,c("model", "DeltaQAICc", "weight", "npar", "QDeviance")] 

## Change parameter names to enhance readability 
## -----------------------------------------------------------------------------
model.table$model <- 
  stringr::str_replace_all(model.table$model, "~", "")
model.table$model <- 
  stringr::str_replace(model.table$model, "Lbinom", "Leucocytozoon")
model.table$model <- 
  stringr::str_replace(model.table$model, "Hatch", "Hatch date")
model.table$model <- 
  stringr::str_replace(model.table$model, "Age_days", "Nestling")
model.table$model <- 
  stringr::str_replace(model.table$model, "BCI", "Body condition")
model.table$model <-
  stringr::str_replace(model.table$model, "ageclass2", "Ageclass2")
model.table$model <- 
  stringr::str_replace(model.table$model, "age", "AgeClass3")
model.table$model <- 
  stringr::str_replace(model.table$model, "Nestling", "Nestling age")
model.table$model <- 
  stringr::str_replace(model.table$model, "y1", "Juvenile")
model.table$model <- 
  stringr::str_replace(model.table$model, "time", "Time")
model.table$model <- 
  stringr::str_replace(model.table$model, "p(1)", "p(.)")
model.table$model <- 
  stringr::str_replace(model.table$model, "F(1)", "F(1)")
model.table$model <- 
  stringr::str_replace(model.table$model, "r(1)", "r(.)")
## Write to csv file: Content is identical to Table S2
readr::write_csv(model.table, "data/adjusted.models.burnham.csv")
```

##### Parameter weights

Compute cumulative weights based on all models including a certain parameter.

```{r}
## Variable importance parameter S
## -----------------------------------------------------------------------------
vars <- c("time", "BCI", "NAO", "Sex", "Rank", "Hatch", "Lbinom", "y1:Age_days", "Morph", "y1:BCI")

S.weights <- sapply(vars, function(var) {
  t <- filter(models.burnham.adj$model.table, stringr::str_detect(S, var))
  sum(t$weight) %>% round(.,3)
})
data.frame(Parameter = vars, weight = S.weights)

## variable importance parameter r
## -----------------------------------------------------------------------------
vars <- c("1", "ageclass2")

r.weights <- sapply(vars, function(var) {
  t <- filter(models.burnham.adj$model.table, stringr::str_detect(r, var))
  sum(t$weight) %>% round(.,3)
})
data.frame(Parameter = vars, weight = r.weights)

## variable importance parameter p
## -----------------------------------------------------------------------------
vars <- c("1", "time")

p.weights <- sapply(vars, function(var) {
  t <- filter(models.burnham.adj$model.table, stringr::str_detect(p, var))
  sum(t$weight) %>% round(.,3)
})
data.frame(Parameter = vars, weight = p.weights)
```


##### Survival rates (S)

Visualise age and time-specific survival estimates. **Note: Except for age and time no other factor plays a role for the survival estimates (e.g., Morph, Rank, infection status)** Hence, rounding estimates at a precision of six digits allows to simplify the output in such a way that there is a single unique estimate per year and age combination.

```{r, message=FALSE, warning=FALSE}
## Format and simplify 
## -----------------------------------------------------------------------------
df.S.age <- lapply(rownames(S.mod.avg), function(row) {
  x <- stringr::str_split(row, " ")[[1]]
  data.frame(parameter = x[1],
             cohort = substr(x[3], 2, nchar(x[3])),
             year = as.numeric(substr(x[5], 2, nchar(x[5]))))
}) %>% 
  do.call("rbind",.) %>% 
  cbind(S.mod.avg) %>% 
  mutate(ageclass = case_when(
    age == 0 ~ 'Juvenile',
    age == 1 ~ 'Subadult',
    age %in% 2:13 ~ 'Adult'  
  )) %>% 
  mutate(ageclass = factor(ageclass, levels = c("Juvenile", "Subadult", "Adult"))) %>% 
  ## round mean, se and CI values at fifth decimal to remove redundant values
  subset(., select = c(year, estimate, se, lcl, ucl, ageclass)) %>% 
  mutate(estimate = round(estimate, 6)) %>% 
  mutate(se = round(se, 6)) %>% 
  mutate(lcl = round(lcl, 6)) %>% 
  mutate(ucl = round(ucl, 6)) %>% 
  unique.data.frame()

## Estimate means for age-classes
## ---------------------------------------------------------------
df.annotation <- df.S.age %>% 
  group_by(ageclass) %>%
  summarise(
    Median = median(estimate),
    mean  = mean(estimate),
    Min = min(estimate),
    Max = max(estimate),
    N = length(estimate),
    SD = stats::sd(estimate)) %>% 
  mutate(SE = SD/sqrt(N))

p.age <- 
  ggplot(df.S.age,
         aes(x = year, y = estimate)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl) ,
                colour = "grey80", size = 1, width = 0) +
  geom_errorbar(aes(ymin = estimate - se, ymax = estimate + se) ,
                colour = "black", size = 1.2, width = 0) +
  geom_line() +
  geom_point(color = "black", fill = "Orange", size = 1.75, shape = 21) +
  theme_article() +
  theme(
    # panel.grid.major.y = element_line(colour = "grey90", size = 0.5),
    # panel.grid.minor.y = element_line(colour = "grey90", size = 0.5),
    axis.line = element_line(size = .7, color = "black"),
    legend.background = element_rect(fill = NA, colour = NA),
    legend.margin = margin(c(1,1,1,1)),
    legend.direction = "horizontal",
    legend.position = "none",
    legend.justification = c(1,1),
    panel.background = element_rect(fill = NA),
    panel.border = element_rect(fill = NA)) +
  scale_color_brewer(palette = "Set1",
                     name = "") +
  scale_y_continuous(breaks = seq(0, 1, .2), limits = c(0, 1)) +
  scale_x_continuous(breaks = seq(2007, 2020, 3), expand = c(0.02,0.02)) +
  xlab("") +
  ylab("Annual survival probability") +
  facet_wrap(~ageclass, ncol = 1, strip.position = "left") +
  geom_text(data = df.annotation, x = 2007, y = -Inf,
            hjust   = 0, vjust   = -0.8, col = "Black", size = 3,
            aes(label = paste0("mean: ",round(mean,2))))
p.age
```

Mean estimates for age and sex combinations over time

```{r, warning=FALSE, results='markup'}
## Age-specific means
## -----------------------------------------------------------------------------
df.S.age %>% 
  group_by(ageclass) %>% 
  summarise(Mean = round(mean(estimate),2),
            Mean.SE = round(sd(estimate)/sqrt(length(estimate)),2),
            Min = round(min(estimate), 2),
            Max = round(max(estimate), 2),
            .groups = "keep") %>% 
  as.data.frame()
```


```{r}
nao$sign <- ifelse(nao$DJFM > 0, "+", "-")
p.nao <- 
    ggplot(filter(nao, year > 2007),
           aes(x = year - 1, y = DJFM, col = sign)) +
    geom_line(col = "black") +
    geom_point(size = 1.75) +
    theme_article() +
    theme(
      legend.position = "none",
      axis.line = element_line(size = .7, color = "black"),
      panel.background = element_rect(fill = NA),
      panel.border = element_rect(fill = NA)) +
    scale_x_continuous(breaks = seq(2007, 2020, 3), expand = c(0.02,0.02)) +
    scale_color_brewer(palette = "Set1",direction = -1) +
    geom_hline(yintercept = 0, lty = "dotted") +
    xlab("Survival period") +
    ylab("NAO")

p.snow <- 
  ggplot(filter(nao, year > 2007),
         aes(x = year - 1, y = snow)) +
  geom_line(col = "black") +
  geom_point(size = 1.75) +
  theme_article() +
  theme(
    legend.position = "none",
    axis.line = element_line(size = .7, color = "black"),
    panel.background = element_rect(fill = NA),
    panel.border = element_rect(fill = NA)) +
  scale_x_continuous(breaks = seq(2007, 2020, 3), expand = c(0.02,0.02)) +
  scale_color_brewer(palette = "Set1",direction = -1) +
  xlab("Survival period") +
  ylab("Snow")

p.temp <- 
  ggplot(filter(nao, year > 2007),
         aes(x = year - 1, y = temp)) +
  geom_line(col = "black") +
  geom_point(size = 1.75) +
  theme_article() +
  theme(
    legend.position = "none",
    axis.line = element_line(size = .7, color = "black"),
    panel.background = element_rect(fill = NA),
    panel.border = element_rect(fill = NA)) +
  scale_x_continuous(breaks = seq(2007, 2020, 3), expand = c(0.02,0.02)) +
  scale_color_brewer(palette = "Set1",direction = -1) +
  xlab("Survival period") +
  ylab("Mean emperature [°C]")

Fig_S8 <- (p.age + theme(axis.text.x = element_blank(),
               axis.ticks.x = element_blank())) +  
  (p.nao + theme(axis.text.x = element_blank(),
                 axis.title.x = element_blank(),
                 axis.ticks.x = element_blank())) +
  p.snow +
  plot_layout(ncol = 1, heights = c(4, 1,1)) +
  plot_annotation(tag_levels = "A")
Fig_S8
```

```{r, echo=FALSE, eval=FALSE}
ggsave(plot = Fig_S8, filename = "plots/Fig_S8.pdf", 
        device = "pdf", dpi = 1200, width = 6, height = 6, units = "in")
ggsave(plot = Fig_S8, filename = "plots/Fig_S8.png", 
        device = "png", dpi = 1200, width = 6, height = 6, units = "in")
```


```{r, eval=FALSE, echo=FALSE}
sessionInfo()
```
