---
title: "Buzzard Survival"
author: "Meinolf Ottensmann"
output: html_document
---

This document details all required steps to repeat analyses presented within our manuscript.

## Load required Packages

```{r, message=FALSE, warning=FALSE}
rm(list = ls())
library(tidyverse)
library(magrittr)
library(patchwork)
library(RMark)
library(sjPlot)
library(sjlabelled)
library(sjmisc)
library(lme4)
library(forestmodel)
```

## Load raw data 
### NAO indices

Here, we use NAO indices that are based on the principal component approach [^8] (Hurrel, 2003).

```{r}
nao <- read.table("data/nao.txt", header = T) %>% 
  mutate(time = as.factor(year))
```

### buzzard data

This analysis is based on data from a longterm study on common buzzards *Buteo buteo* comprising 2499 individuals that were ringed as nestlings and individually marked with wingtags in the period 2007 to 2019. Note, due to logistic constraints there is no information on *Leucocytozoon* infections available for the cohort of 2008 and thus individuals from that cohort are omitted from analyses.
There are two datasets, `ringing_data` containing morphometric measures of all individuals used for this analysis and `resighting_data` detailing resightings for these individuals.

```{r, message=FALSE}
ringing_data <- as.data.frame(read_csv("data/ringing_data.csv")) 
resighting_data <- as.data.frame(read_csv("data/resighting_data.csv"))
```

# Statistical Analysis

## Burnham model (Live & dead encounters)

We model survival probabilites using combined dead and live encounter events using the `Burnham` model in `MARK`.  

Based on ringing data and resightings of individuals, we assemble combined live-dead encounter histories for all individuals using the function `create.ld`. This functions generates a vector of the form `LDLDLD`, where `L` codes live encounters and `D` dead recoveries respectively. Together, each `LD` pair codes for one interval of the study period.

```{r}
## function to derive encounter histories
create.ld <- function(start, end, ringing_data, resights) {
  ## Identify all individuals 
  individuals <- ringing_data$Ring %>%
    unique()
  
  ## define Study periods based on start and end years
  years <- c(start, end)
  periods <- rep(seq(start, end), each = 1)

  ## create matrix 
  ## rows: Individuals
  ## cols: 2 Periods per year with Live and Dead encounters respectively
  mat <- matrix(0, nrow = length(individuals), ncol = length(periods)*2) %>% 
    set_colnames(paste0(rep(periods, each = 2), c("L", "D"))) %>% 
    set_rownames(individuals)
  
  ## code year of capture as 1,
  ## Code sightings in subsequent intervals by 1 too
  for (x in individuals) {
    year.ringed <- filter(ringing_data, Ring == x)[["Year"]][1]
      ## code birth year in living encounters
      mat[rownames(mat) == x, colnames(mat) == paste0(year.ringed, "L")] <- 1  

      ## get resights and assign to sampling instances. As years are defined
      ## by the period 01.04 to 31.03, the first three month of each year count 
      ## for the preceding interval.
      year.resighted <- 
        filter(resights, Ring == x)[,c("Year", "Month")] %>% 
        mutate(period = case_when(
          Month == "01" ~ paste0(as.numeric(Year) - 1 , "L"),
          Month == "02" ~ paste0(as.numeric(Year) - 1 , "L"),
          Month == "03" ~ paste0(as.numeric(Year) - 1 , "L"),
          Month == "04" ~ paste0(Year, "L"),
          Month == "05" ~ paste0(Year, "L"),
          Month == "06" ~ paste0(Year, "L"),
          Month == "07" ~ paste0(Year, "L"),
          Month == "08" ~ paste0(Year, "L"),
          Month == "09" ~ paste0(Year, "L"),
          Month == "10" ~ paste0(Year, "L"),
          Month == "11" ~ paste0(Year, "L"),
          Month == "12" ~ paste0(Year, "L")
        ))

        ## write resights to matrix 
  for (col in year.resighted$period) {
    mat[rownames(mat) == x, colnames(mat) == col] <- 1
  }
  
  ## check if reported as dead
  death <- filter(ringing_data, Ring == x)[,c("Dead", "DateDeath")][1,]
  if (!is.na(death$Dead)) {
    col <- paste0(substr(death$DateDeath, 1, 4),"D")
    mat[rownames(mat) == x, colnames(mat) == col] <- 1
  }
  
}

## coerce to character vector
return(apply(mat[,], 1, function(x) paste0(x, collapse = "")) %>% 
         as.character())

}
```

Apply the function to the data:

```{r}
## add resighting histories to the ringing data
df <- data.frame(
  filter(ringing_data, !is.na(Lprev)),
  ch =  create.ld(start = 2007, # first year
                  end = 2020, # last year
                  ringing_data = ringing_data,
                  resights = resighting_data)) %>% 
  mutate(ch = as.character(ch)) %>% 
  mutate(Sex = as.factor(Sex)) %>% 
  mutate(Hatch = scale(Hatch)) 

## process the data for RMark
dp  <-  process.data(df, model = "Burnham",
                     begin.time = 2007, groups = "Sex")

## create design data
ddl <-  make.design.data(dp)

## Add design data for components 
## Survival
ddl$S <- ddl$S %>% 
  mutate(age = as.numeric(as.character(age))) %>% 
  mutate(ageclass = 
           case_when(
             age == 0 ~ "CY1",
             age == 1 ~ "CY2",
             age == 2 ~ "CY3",
             age >= 3 ~ "CY4plus"
           ))

## collect variables which are year but not individual specific
covars <- left_join(data.frame(time = ddl$S$time), nao, "time")

## extend design data with covariates
ddl$S$djfm <- covars$djfm ## Winter NAO index

## Site-Fidelity
ddl$F <- ddl$F %>% 
  mutate(age = as.numeric(as.character(age))) %>% 
  mutate(ageclass = 
           case_when(
             age == 0 ~ "CY1",
             age == 1 ~ "CY2",
             age == 2 ~ "CY3",
             age >= 3 ~ "CY4plus"
           ))

## recovery chances
ddl$r <- ddl$r %>%
  mutate(age = as.numeric(as.character(age))) %>% 
  mutate(ageclass = 
           case_when(
             age == 0 ~ "CY1",
             age == 1 ~ "CY2",
             age == 2 ~ "CY3",
             age >= 3 ~ "CY4plus"
           ))

## resight probability
ddl$p <- ddl$p %>% 
  mutate(age = as.numeric(as.character(age))) %>% 
  mutate(ageclass = 
           case_when(
             age == 0 ~ "CY1",
             age == 1 ~ "CY2",
             age == 2 ~ "CY3",
             age >= 3 ~ "CY4plus"
           ))
## Goodness of fit using RELEASE
release.gof(dp)
rm(covars); rm(nao); 
```

```{r, eval=FALSE, echo=FALSE}
dat <- sapply(df$ch, stringr::str_split, "", simplify = T) %>% 
  do.call("rbind",.) %>% 
   apply(.,2,as.numeric) %>% 
  set_colnames(rep(c("L", "D"), nchar(df$ch[[1]])/2)) 

## split dead and live
L <- dat[,seq(1, ncol(dat), 2)] %>% 
  set_colnames(2007:2020)
D <- dat[,seq(2, ncol(dat), 2)] %>% 
  set_colnames(2007:2020)

rm(dat)

output <- matrix(
  data = 0,
  nrow = length(seq(2007, 2020)),
  ncol = 2 + length(seq(2008, 2020)) + length(seq(2007, 2020))
) %>% 
  as.data.frame()
colnames(output) <- 
  c("Year", "Releases", 
    paste0("L", 2008:2020),
    paste0("D", 2007:2020))
output[1:nrow(output),1] <- 2007:2020

r <- 1
for (r in 1:nrow(L)) {
  alive <- names(which(L[r,] == 1)) 
  released <- min(alive)
  resighted <- alive[alive != released]
  dead <- ifelse(any(D[r,] == 1), names(which(D[r,] == 1)),character())
  
  ## add release to table
  output[output$Year == released, "Releases"] <-
      output[output$Year == released, "Releases"] + 1
  
  if (length(resighted > 0)) {
    output[output$Year == released, paste0("L", resighted)] <-
      output[output$Year == released, paste0("L", resighted)] + 1
  }
  if (length(dead[!is.na(dead)] > 0)) {
    output[output$Year == released, paste0("D", dead)] <-
      output[output$Year == released, paste0("D", dead)] + 1
  }
}
write.csv(output, file = "data/sample.sizes.csv")
knitr::kable(output)
```

```{r,eval=F}
resights <- sapply(df.A$ch, stringr::str_split, "", simplify = T) %>% 
  do.call("rbind",.) %>% 
  apply(.,2,as.numeric) %>%
  set_colnames(2007:2019) %>% 
  colSums() 
resights <- data.frame(Year = names(resights),
                       Resights = resights)

combo <- left_join(resights, cohorts) %>% 
  mutate(Resights = ifelse(is.na(Resights), 0, Resights)) %>% 
  mutate(Freq = ifelse(is.na(Freq), 0, Freq)) %>% 
  mutate(annual.resights = Resights - Freq)
write.csv(combo,"data/sample.sizes.csv")


## resighs per ind
res.ind <- sapply(df.A$ch, stringr::str_split, "", simplify = T) %>% 
  do.call("rbind",.) %>% 
  apply(.,2,as.numeric) %>%
  set_colnames(2007:2019) %>% 
  rowSums() - 1
cat("Ind Resighted:", length(res.ind[res.ind > 0]))
cat("Prop Resighted:", length(res.ind[res.ind > 0])/length(res.ind))
cat("Avg #:", mean(res.ind[res.ind > 0]))
cat("Avg #:", sd(res.ind[res.ind > 0]))

## format sample size report
cohorts <- data.frame(with(df.B, table(Year)))
sex <- data.frame(with(df.B, table(Year, Sex)))
leuco <- with(df.B, table(Lprev)) # summary by leucocytozoon count

resights <- sapply(df.B$ch, stringr::str_split, "", simplify = T) %>% 
  do.call("rbind",.) %>% 
  apply(.,2,as.numeric) %>%
  #set_colnames(2007:2019) %>% 
  colSums() 
resights <- data.frame(Year = names(resights),
                       Resights = resights)

combo <- left_join(resights, cohorts) %>% 
  mutate(Resights = ifelse(is.na(Resights), 0, Resights)) %>% 
  mutate(Freq = ifelse(is.na(Freq), 0, Freq)) %>% 
  mutate(annual.resights = Resights - Freq)
write.csv(combo,"data/sample.sizes.csv")


## resighs per ind
res.ind <- sapply(df.B$ch, stringr::str_split, "", simplify = T) %>% 
  do.call("rbind",.) %>% 
  apply(.,2,as.numeric) %>%
  set_colnames(2007:2019) %>% 
  rowSums() - 1
cat("Ind Resighted:", length(res.ind[res.ind > 0]))
cat("Prop Resighted:", length(res.ind[res.ind > 0])/length(res.ind))
cat("Avg #:", mean(res.ind[res.ind > 0]))
cat("Avg #:", sd(res.ind[res.ind > 0]))

```

### Bootstrapping Global model

The code below is used to implement a parametric bootstrapping to calculate median $\hat{C}$ based in function defined in the script `R/bootstrapping_functions.R`

```{r, eval = F, message=FALSE, warning=FALSE}
source("R/bootstrapping.R")
## run global model to get parameter estimates
global.model <- mark(dp, ddl, silent = T, output = F,
                     model.parameters = list(
S = list(formula = ~ time*Sex), 
p = list(formula = ~ time*Sex),
F = list(formula = ~ time*Sex),
r = list(formula = ~ time*Sex)),
model = 'Burnham')

## get model estimates/effects
params <- global.model$results$real

## get parameter names
components <- lapply(row.names(params), function(x) {
  stringr::str_split(x, pattern = " ")[[1]]
}) %>% 
  do.call("rbind",.) %>% 
  as.data.frame() %>% 
  set_colnames(c("Estimate", "Group", "Cohort", "a", "t"))

## extract sex-specific components
S.female <-   
  params$estimate[which(components$Estimate == "S" & components$Group == "gFemale")]
S.male <-
  params$estimate[which(components$Estimate == "S" & components$Group == "gMale")] 
p.female <-
  params$estimate[which(components$Estimate == "p" & components$Group == "gFemale")]
p.male <-
  params$estimate[which(components$Estimate == "p" & components$Group == "gMale")]
F.female <-
  params$estimate[which(components$Estimate == "F" & components$Group == "gFemale")]
F.male <-
  params$estimate[which(components$Estimate == "F" & components$Group == "gMale")]
r.male <-
  params$estimate[which(components$Estimate == "r" & components$Group == "gMale")]
r.female <-
  params$estimate[which(components$Estimate == "r" & components$Group == "gFemale")]

## conduct simulations using parametric bs for both models
sim.out <- sims.burnham(
  S.fem = S.female, 
  S.mal = S.male,
  p.fem = p.female,
  p.mal = p.male,
  F.fem = F.female,
  F.mal = F.male,
  r.fem = r.female,
  r.mal = r.male,
  marked =  Marked.burnham(data = df,
                   n.occasions = nchar(df$ch[1])/2,
                   groups = c("Female","Male"),
                   group.col = "Sex"),
  group.col = "Sex", reps = 9, 
  model.parameters = list(
S = list(formula = ~ time*sex), 
p = list(formula = ~ time*sex),
F = list(formula = ~ time*sex),
r = list(formula = ~ time*sex)))

save(sim.out, file = "data/sim.out.RData")
c.hat <- global.model$results$deviance/sim.out$deviance.mean
save(c.hat, file = "data/c.hat.RData")
cleanup(ask = F)
```

### Candidate models

The code below implements all candidate models. As the execution is computationally extensive the code is not executed by default but the output is saved as part of this repository.  

```{r, eval=FALSE, message=FALSE, warning=FALSE}
do_analysis = function(dp, ddl) {
  
  S.dot = list(formula = ~ 1)
  S.0 = list(formula = ~ ageclass + Sex)
  S.cond = list(formula = ~ ageclass*cond + Sex)
  S.condxdjfm = list(formula = ~ ageclass*cond + ageclass*djfm + Sex)
  S.condxdjfmxhatch = list(formula = ~ ageclass*cond + ageclass*djfm + ageclass*Hatch + Sex)
  S.djfm = list(formula = ~ ageclass*djfm + Sex)
  S.djfmxhatch = list(formula = ~ ageclass*djfm + ageclass*Hatch + Sex)
  
  S.A = list(formula = ~ 
               ageclass * cond + ageclass * djfm + Sex + Morph + ageclass*Lprev + ageclass * Ekto)
  
  S.B = list(formula = ~ 
               ageclass * cond + ageclass * djfm + Sex + Morph + ageclass*Lprev + ageclass * Ekto + ageclass * ageclass * Hatch + ageclass * Rank)
  
  S.C = list(formula = ~ 
               ageclass * cond + ageclass * djfm + Sex + Morph + ageclass*Lprev + ageclass * Ekto + ageclass * ageclass * Hatch + ageclass * Rank * Brood_Size)
  
  
  S.D = list(formula = ~ 
               ageclass * cond + ageclass * djfm + ageclass*Lprev)
  
  
  S.E = list(formula = ~ 
               ageclass * cond + djfm + ageclass*Lprev)
  
  
  p.time = list(formula = ~ time + ageclass)
  p.dot = list(formula = ~ 1)
  F.age = list(formula = ~ Sex + ageclass)
  F.age2 = list(formula = ~ Sex * ageclass)
  F.dot = list(formula = ~ 1)
  F.sex = list(formuula = ~ Sex)
  r.age = list(formula = ~ ageclass)
  r.dot = list(formula = ~ 1)
  
  cml = create.model.list(model = "Burnham")
  # run all models and return as a list with class 'marklist'
  results = mark.wrapper(cml, data = dp, ddl = ddl, output = F, silent = T)
  return(results)
}
models <- do_analysis(dp = dp, ddl = ddl)

## adjust for overdispersion using c-hat
load("data/c.hat.RData")
models.adj <-  adjust.chat(model.list = models, chat = c.hat) 
model.tab <- model.table(models.adj, use.lnl = TRUE) %>% 
  mutate(weight = round(weight, 2)) %>% 
  mutate(DeltaQAICc = round(DeltaQAICc, 2)) %>% 
  mutate(Neg2LnL = round(Neg2LnL, 2)) 

# models within 7 DeltaQAICc units of the best are considered to have
## statistical support
supported.models <- remove.mark(
  marklist = models.adj,
  model.numbers = which(model.table(models.adj, sort = F)[["DeltaQAICc"]] > 7))

suffix <- Sys.time() %>% 
  as.character() %>% 
  stringr::str_remove(., " ") %>% 
  gsub(pattern = ":", replacement = "-", x = .)

print(model.tab[,c("model", "npar", "DeltaQAICc", "Neg2LnL", "weight")])
write.csv(model.tab, "data/model_output.csv", row.names = F)

save(suffix, file = paste0("data/",suffix,".RData"))
save(models, file = paste0("data/models.", suffix, ".RData"))
save(models.adj, file = paste0("data/models.adj.", suffix, ".RData"))
save(supported.models, file = paste0("data/supported.models.", suffix, ".RData"))
```

### Results

As the execution of the models is time consuming, we provide the captured output as part of the repository 

```{r}
## load previously saved results
#suffix <- '2020-10-0522-27-47'
suffix <- '2020-10-0820-59-18'

load(paste0("data/models.", suffix, ".RData"))
load(paste0("data/models.adj.", suffix, ".RData"))
load(paste0("data/supported.models.", suffix, ".RData"))
```

#### Model averaging

```{r, message=FALSE, warning=FALSE}
## Model average Survival estimates (S) 
S.mod.avg <- model.average(supported.models, parameter = "S", drop = T, vcv = T)[["estimates"]] %>% 
  subset(., select = c(estimate, se, lcl, ucl, age, Sex))

## Model average Fidelity estimates (F)
F.mod.avg <- model.average(models.adj, parameter = "F", vcv = T, drop = T)$estimates %>% 
  subset(., select = (c(estimate, se, lcl, ucl, time, age, Sex))) %>% 
  unique.data.frame() 

## Model average recovery estimates (r)
r.mod.avg <- model.average(models.adj, parameter = "r", vcv = T, drop = T)$estimates %>% 
  subset(., select = (c(estimate, se, lcl, ucl, time, age, Sex))) %>% 
  unique.data.frame() 

## Model average resighting estimates (p)
p.mod.avg <- model.average(models.adj, parameter = "p", vcv = T, drop = T)$estimates %>% 
  subset(., select = (c(estimate, se, lcl, ucl, time, age))) %>% 
  unique.data.frame() 
```

#### Survival probabilites

##### Age

```{r, message=FALSE, warning=FALSE}
## Compute means for sex by age combinations
comb <- list(list(Sex = "Male", age = 0, ageclass = "CY1"), 
             list(Sex = "Male", age = 1, ageclass = "CY2"),
             list(Sex = "Male", age = 2, ageclass = "CY3"),
             list(Sex = "Male", age = 3:13, ageclass = "CY4plus"),
             list(Sex = "Female", age = 0, ageclass = "CY1"), 
             list(Sex = "Female", age = 1, ageclass = "CY2"),
             list(Sex = "Female", age = 2, ageclass = "CY3"),
             list(Sex = "Female", age = 3:13, ageclass = "CY4plus")
             )
df.S.age <- 
lapply(1:length(comb), function(x) {
  dx <- filter(S.mod.avg, Sex == comb[[x]]$Sex, age %in% comb[[x]]$age)
  data.frame(estimate = mean(dx$estimate),
             lcl = mean(dx$lcl), 
             ucl = mean(dx$ucl),
             se = mean(dx$se),
             Sex = comb[[x]]$Sex,
             Age = comb[[x]]$ageclass)
}) %>% 
  do.call("rbind",.) %>% 
  mutate(x = c(0.8, 1.8, 2.8, 3.8, 1.2, 2.2, 3.2, 4.2))

p.age <- 
  ggplot(df.S.age, aes(x = x, y = estimate, col = Sex, fill = Sex)) +
  geom_errorbar(aes(ymin = estimate - se, ymax = estimate + se), width = .25, size = 1) +
  geom_point(size = 3, shape = 15) +
  theme_classic() +
  theme(
    panel.border = element_rect(fill = NA),
    panel.background = element_rect(fill = "Grey95"),
    text = element_text(size = 12),
    legend.position = c(.2, .05),
    legend.title = element_blank(),
    legend.direction = "horizontal",
    legend.background = element_rect(fill = NA),
    aspect.ratio = 1) +
  scale_colour_brewer(palette = "Set1", direction = -1) +
  scale_x_continuous(breaks = 1:4, labels = unique(df.S.age$Age)) +
  scale_y_continuous(expand = c(0,0), 
                     limits = c(0,1),
                     breaks = seq(0, 1, .2)) +
  xlab("Age class") +
  ylab("Suvival probability [Mean \U00B1 SE]") 

p.age
```


##### Condition

```{r message=FALSE, warning=FALSE}
# create values of cond.index to use for predictions
cond.index.values = seq(from = -2, to = 2, length.out = 100)

## Extract get 'par.index' of interest
index.meta <- data.frame(
  Sex = rep(c("Female", "Male"), each = 4),
  ageclass = rep(c("CY1", "CY2", "CY3", "CY4plus"), 2)) %>% 
  ## add par.index
  left_join(., ddl$S[,c("par.index", "Sex", "ageclass")],
            by = c("Sex", "ageclass")) %>% 
  ## code additional vars
  mutate(dummy = paste0(Sex, ageclass)) 

## simplify, one par.index for each dummy var
model.indices <- sapply(unique(index.meta$dummy), function(x) {
  ## return first row matching the query
  index.meta[index.meta$dummy == x,"par.index"][1]
}) %>%
  unlist()

## subset index.meta
index.meta <- index.meta[index.meta$par.index %in% model.indices,]

## get estimates from the model
pred.model <- covariate.predictions(
  supported.models, 
  data = data.frame(
    cond = cond.index.values), 
  indices = model.indices %>% as.numeric(),
  drop = T)[["estimates"]]

## extend df with grouping variables
pred.model <- left_join(pred.model,
                        index.meta[,c("par.index", "ageclass", "Sex")], 
                        by = "par.index") %>%
  mutate(Group = paste0(Sex, ageclass))

p.cond <- 
  ggplot(pred.model, aes(x = covdata, y = estimate, group = Group)) +
  geom_line(size = 1.5,aes(colour = Sex)) +
  geom_ribbon(aes(ymin = estimate - se,
                  ymax = estimate + se,
                  colour = Sex), alpha = 0.2) +
  scale_colour_brewer(palette = "Set1") + 
  theme_classic() +
  theme(
    panel.border = element_rect(fill = NA),
    panel.background = element_rect(fill = "Grey95"),
    text = element_text(size = 12),
    legend.position = 'none',
    aspect.ratio = 1) +
  xlab("Standardised Body condition") +
  ylab("Survival probability [mean \U00B1 SE]") +
  scale_y_continuous(expand = c(0,0), limits = c(0,1),
                     breaks = seq(0, 1, 0.2)) +
  facet_wrap(~ageclass)
p.cond
```

##### NAO

```{r, message=FALSE, warning=FALSE}
djfm.values = seq(from = min(ddl$S$djfm),
                  to = max(ddl$S$djfm),
                  length.out = 10)

## Extract get par.index of interest
index.meta <- data.frame(
  ageclass = rep(c("CY1", "CY2", "CY3", "CY4plus"), 1)) %>%   
  ## add par.index
  left_join(., ddl$S[,c("par.index", "ageclass", "djfm")]) %>% 
  ## code additional vars
  mutate(dummy = paste0(djfm, ageclass)) 

## simplify, one par.index for each dummy var.
model.indices <- sapply(unique(index.meta$dummy), function(x) {
  ## return first row matching the query
  index.meta[index.meta$dummy == x,"par.index"][1]
}) %>%
  unlist()

index.meta <- index.meta[index.meta$par.index %in% model.indices,]

pred.djfm <- covariate.predictions(
  supported.models, 
  data = data.frame(
    djfm = djfm.values), 
  indices = index.meta$par.index)[["estimates"]]

## extend df with grouping variables
pred.model <- left_join(
  pred.djfm,
  index.meta[,c("par.index", "ageclass")],
  by = "par.index") %>% 
  mutate(Group = ageclass)

pred.model <- left_join(
  index.meta,
  pred.model[,c("par.index", "estimate", "lcl", "ucl", "se", "ageclass", "Group")]) %>% 
  unique.data.frame()

p.djfm <- 
  ggplot(pred.model, aes(x = djfm, y = estimate, group = Group)) +
  geom_line(size = 1.5) +
  geom_ribbon(aes(ymin = estimate - se,
                  ymax = estimate + se),
              alpha = 0.2) +
  scale_colour_brewer(palette = "Set1") + 
  theme_classic() +
  theme(legend.position = 'none',
        panel.border = element_rect(fill = NA),
        panel.background = element_rect(fill = "Grey95"),
        text = element_text(size = 12),
        aspect.ratio = 1) +
  xlab("DJFM NAO") +
  ylab("Survival probability [mean \U00B1 CI]") +
  facet_wrap(~ageclass)

p.djfm
```

##### Leucocytozoon

```{r, message=FALSE, warning=FALSE}
leuco = 0:4

## Extract get par.index of interest
index.meta <- data.frame(
  ageclass = rep(c("CY1", "CY2", "CY3", "CY4plus"), 1)) %>%   
  ## add par.index
  left_join(., ddl$S[,c("par.index", "ageclass")]) %>% 
  ## code additional vars
  mutate(dummy = paste0(ageclass)) 

## simplify, one par.index for each dummy var
model.indices <- sapply(unique(index.meta$dummy), function(x) {
  ## return first row matching the query
  index.meta[index.meta$dummy == x,"par.index"][1]
}) %>% 
  unlist()

index.meta <- index.meta[index.meta$par.index %in% model.indices,]

pred.leuco <- covariate.predictions(
  supported.models, 
  data = data.frame(
    Lprev = leuco, bl.index.sq = leuco^2), 
  indices = index.meta$par.index)[["estimates"]]

## extend df with grouping variables
pred.model <- left_join(pred.leuco, index.meta[,c("par.index", "ageclass")], by = "par.index") %>% 
  mutate(Group = ageclass)

pred.model <- left_join(index.meta, pred.model[,c("par.index", "Lprev", "estimate", "lcl", "ucl", "se", "ageclass", "Group")]) %>% 
  unique.data.frame()

p.leuco <- 
  ggplot(pred.model, aes(x = Lprev, y = estimate, group = ageclass)) +
  geom_errorbar(aes(ymin = estimate - se, ymax = estimate + se), width = .25, size = 1.5) +
  geom_point(size = 3, colour = "Black", shape = 15) +
  scale_colour_brewer(palette = "Set1") + 
  theme_classic() +
  theme(
    panel.border = element_rect(fill = NA),
    panel.background = element_rect(fill = "Grey95"),
    text = element_text(size = 12),
    legend.position = 'bottom',
    aspect.ratio = 1) +
  xlab("Leucozytozoon infection intensity") +
  ylab("Survival probability [mean \U00B1 SE]") +
  facet_wrap(~ageclass)

p.leuco
```

##### Carnus 

```{r, message=FALSE, warning=FALSE}
ekto = 0:4

## Extract get par.index of interest
index.meta <- data.frame(
  ageclass = rep(c("CY1", "CY2", "CY3", "CY4plus"), 1)) %>%   
  ## add par.index
  left_join(., ddl$S[,c("par.index", "ageclass")]) %>% 
  ## code additional vars
  mutate(dummy = paste0(ageclass)) 

## simplify, one par.index for each dummy var
model.indices <- sapply(unique(index.meta$dummy), function(x) {
  ## return first row matching the query
  index.meta[index.meta$dummy == x,"par.index"][1]
}) %>% 
  unlist()

index.meta <- index.meta[index.meta$par.index %in% model.indices,]

pred.ekto <- covariate.predictions(
  supported.models, 
  data = data.frame(
    Ekto = ekto), 
  indices = index.meta$par.index)[["estimates"]]

## extend df with grouping variables
pred.model <- left_join(pred.ekto, index.meta[,c("par.index", "ageclass")], by = "par.index") %>% 
  mutate(Group = ageclass)

pred.model <- left_join(index.meta, pred.model[,c("par.index", "covdata", "estimate", "lcl", "ucl", "ageclass", "Group")]) %>% 
  unique.data.frame()

p.ekto <- 
  ggplot(pred.model, aes(x = covdata, y = estimate, group = ageclass)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), width = .25, size = 1.5) +
  geom_point(size = 3, colour = "Black", shape = 15) +
  scale_colour_brewer(palette = "Set1") + 
  theme_classic() +
  theme(
    panel.border = element_rect(fill = NA),
    panel.background = element_rect(fill = "Grey95"),
    text = element_text(size = 12),
    legend.position = 'bottom',
    aspect.ratio = 1) +
  xlab("Ectoparasite infection intensity") +
  ylab("Survival probability [mean \U00B1 CI]") +
  facet_wrap(~ageclass)

p.ekto
```

#### Resighting probabilies

All supported models contain a time dependent effect on resighting probabilties, with an average estimate of `r mean(p.mod.avg$estimate)` (range: `r min(p.mod.avg$estimate)` - `r max(p.mod.avg$estimate)`) .

```{r, message=FALSE, warning=FALSE}
## compute means for sex and age combinations
## exclude fixed values for know
comb <- list(
  list(time = 2008, age = 1, Age = "CY2"),
 
  list(time = 2009, age = 1, Age = "CY2"),
  list(time = 2009, age = 2, Age = "CY3"),
  
  list(time = 2010, age = 1, Age = "CY2"),
  list(time = 2010, age = 2, Age = "CY3"),
  list(time = 2010, age = 3:13, Age = "CY4+"),           
  list(time = 2011, age = 1, Age = "CY2"),
  list(time = 2011, age = 2, Age = "CY3"),
  list(time = 2011, age = 3:13, Age = "CY4+"),           
  list(time = 2012, age = 1, Age = "CY2"),
  list(time = 2012, age = 2, Age = "CY3"),
  list(time = 2012, age = 3:13, Age = "CY4+"),         
  list(time = 2013, age = 1, Age = "CY2"),
  list(time = 2013, age = 2, Age = "CY3"),
  list(time = 2013, age = 3:13, Age = "CY4+"),         
  list(time = 2014, age = 1, Age = "CY2"),
  list(time = 2014, age = 2, Age = "CY3"),
  list(time = 2014, age = 3:13, Age = "CY4+"),         
  list(time = 2015, age = 1, Age = "CY2"),
  list(time = 2015, age = 2, Age = "CY3"),
  list(time = 2015, age = 3:13, Age = "CY4+"),         
  list(time = 2016, age = 1, Age = "CY2"),
  list(time = 2016, age = 2, Age = "CY3"),
  list(time = 2016, age = 3:13, Age = "CY4+"),         
  list(time = 2017, age = 1, Age = "CY2"),
  list(time = 2017, age = 2, Age = "CY3"),
  list(time = 2017, age = 3:13, Age = "CY4+"),         
  list(time = 2018, age = 1, Age = "CY2"),
  list(time = 2018, age = 2, Age = "CY3"),
  list(time = 2018, age = 3:13, Age = "CY4+"),         
  list(time = 2019, age = 1, Age = "CY2"),
  list(time = 2019, age = 2, Age = "CY3"),
  list(time = 2019, age = 3:13, Age = "CY4+"),         
  list(time = 2020, age = 1, Age = "CY2"),
  list(time = 2020, age = 2, Age = "CY3"),
  list(time = 2020, age = 3:13, Age = "CY4+")     
)
p.effect <- 
lapply(1:length(comb), function(x) {
  dx <- filter(p.mod.avg,
               time == comb[[x]]$time,
               age %in% comb[[x]]$age)
  data.frame(estimate = mean(dx$estimate),
             lcl = mean(dx$lcl), 
             ucl = mean(dx$ucl),
             se = mean(dx$se),
             time = comb[[x]]$time,
             Age = comb[[x]]$Age)
}) %>% 
  do.call("rbind",.) %>% 
  mutate(x = case_when(
    Age == "CY2" ~ time - 0.25,
    Age == "CY3" ~ time,
    Age == "CY4+" ~ time + 0.25
  ))


p.res <- ggplot(p.effect, aes(x = x, 
                              y = estimate,
                              col = Age)) +
  theme_classic() +
  theme(
    panel.border = element_rect(fill = NA),
    panel.background = element_rect(fill = "Grey95"),
    text = element_text(size = 12)) +
  geom_errorbar(aes(ymin = estimate - se, ymax = estimate + se), width = .25, size = 1.5) +
  geom_point(size = 1.5, shape = 15) +
  xlab("Year") +
  ylab("Resighting probability [mean \U00B1 SE]") +
  scale_x_continuous(breaks = seq(2008, 2020, 3)) + 
  scale_y_continuous(expand = c(0,0), 
                     limits = c(0,1),
                     breaks = seq(0, 1, .2)) 
p.res
```

#### Recovery probabilites

```{r, eval=F}
mod.avg.r <- 
lapply(1:length(comb), function(x) {
  dx <- filter(r.mod.avg,
               #Sex == comb[[x]]$Sex,
               age %in% comb[[x]]$age)
  data.frame(estimate = mean(dx$estimate),
             lcl = mean(dx$lcl), 
             ucl = mean(dx$ucl), 
             se = mean(dx$se),
             #Sex = comb[[x]]$Sex,
             Age = comb[[x]]$Age)
}) %>% 
  do.call("rbind",.) 
# %>% 
#   mutate(x = c(0.8, 1.8, 2.8, 3.8, 1.2, 2.2, 3.2, 4.2))
 
p.recov <- ggplot(mod.avg.r, aes(x = x, y = estimate, col = Sex)) +
  theme_classic() +
  theme(legend.position = "none",
        legend.background = element_rect(fill = NA),
        panel.border = element_rect(fill = NA),
        panel.background = element_rect(fill = "Grey95"),
        text = element_text(size = 12)) +
  geom_errorbar(aes(ymin = estimate - se, ymax = estimate + se), width = .25, size = 1.5) +
  geom_point(size = 3, shape = 15) +
  xlab("Age class") +
  ylab("Recovery chances [mean \U00B1 SE]") +
  scale_x_continuous(breaks = c(1:4), labels = c("CY1", "CY2", "CY3", "CY4plus")) +
  scale_y_continuous(expand = c(0,0),
                     limits = c(-0.2,1.2),
                     breaks = seq(0, 1, .2)) +
  scale_colour_brewer(palette = "Set1")

p.recov
```



#### Fidelity (F)

```{r, message=FALSE, warning=FALSE}
df.F <- 
lapply(1:length(comb), function(x) {
  dx <- filter(F.mod.avg, Sex == comb[[x]]$Sex, age %in% comb[[x]]$age)
  data.frame(estimate = mean(dx$estimate),
             lcl = mean(dx$lcl), 
             ucl = mean(dx$ucl), 
             se = mean(dx$se),
             Sex = comb[[x]]$Sex,
             Age = comb[[x]]$ageclass)
}) %>% 
  do.call("rbind",.) %>% 
  mutate(x = c(0.8, 1.8, 2.8, 3.8, 1.2, 2.2, 3.2, 4.2))
 
p.fidel <- ggplot(df.F, aes(x = x, y = estimate, col = Sex)) +
  theme_classic() +
  theme(legend.position = "none",
        legend.background = element_rect(fill = NA),
        panel.border = element_rect(fill = NA),
        panel.background = element_rect(fill = "Grey95"),
        text = element_text(size = 12)) +
  geom_errorbar(aes(ymin = estimate - se, ymax = estimate + se), width = .25, size = 1.5) +
  geom_point(size = 3, shape = 15) +
  xlab("Age class") +
  ylab("Site fidelity [mean \U00B1 SE]") +
  scale_x_continuous(breaks = c(1:4), labels = c("CY1", "CY2", "CY3", "CY4plus")) +
  scale_y_continuous(expand = c(0,0),
                     limits = c(0,1.2),
                     breaks = seq(0, 1, .2)) +
  scale_colour_brewer(palette = "Set1")

p.fidel
```


# D.) Known cases of first year survival

```{r}
## identify all cases where individuals died during the first year of life
dead <-
  filter(ringing_data,
         Dead == 'Yes',
         !CauseDeath %in% c('Jumper', 'Poisoned', 'Shot')) %>% 
  mutate(HatchingDate = as.Date(paste0(Year, "-01-01")) + Hatch) %>% 
  mutate(DaysSinceHatching = as.Date(DateDeath) - HatchingDate) %>% 
    filter(., DaysSinceHatching <= 366)

dead <- dead$Ring

## Identify all cases where individuals where alive after completing the first
## year of life, i.e. were resighted
alive <- resighting_data %>% 
  filter(DaysSinceHatching > 365)
alive <- alive$Ring %>% unique()

## merge to dataframe
df <- data.frame(Ring = c(dead, alive), 
                 Group = c(rep("Dead", length(dead)), rep("Alive", length(alive)))) %>% 
  left_join(., ringing_data, by = "Ring") %>% 
  mutate(surv = ifelse(Group == "Dead", 0, 1)) %>% 
  mutate(Leucocytozoon = ifelse(Lbinom == 0, "Uninfected", "Infected")) %>% 
  mutate(Carnus = ifelse(Ekto == 0, "Uninfected", "Infected")) %>% 
  mutate(Morph = case_when(
    Morph == "2" ~ "Dark",
    Morph == "3" ~ "Intermediate",
    Morph == "4" ~ "Light")) %>% 
  mutate(Hatch = scale(Hatch)) %>% 
  mutate(Age = scale(Age_days))

model <- glmer(
  surv ~ Leucocytozoon + cond + Sex + Carnus + Hatch + Age + (1|Year) ,
               data = df,
               family = "binomial")

model.plot <- plot_model(model = model, sort.est = TRUE, transform = NULL,
                         
                         
                         
           title = "",
           ci.lvl = .95,
           ci.style = "whisker",
           dot.size = 3,
           line.size = 1.2,
           ) +
  theme_classic(base_size = 12) +
  geom_hline(yintercept = 0, lty = "dashed") +
  ylab("Effect size") + xlab("Model parameter") +
  ggtitle(label = "First year survival", 
          subtitle = paste0(" (Survived: ", length(df$surv[df$surv == 1]),
                            " Died: ", length(df$surv[df$surv == 0]), ")")) 
model.plot
tab_model(model,
          show.stat = T,
          string.est = "Estimate",
          string.ci = "95% CI",
          transform = NULL,
          collapse.ci = F)
```


#### Acknowledgements

The implemenation for calculating the adjusted $\hat C$ via bootstrapping was based on a modified R code created by Michael J. Conroy (accessed from https://sites.google.com/site/cmrsoftware/lecture-lab-schedule/6---model-fit-and-mmi/methods-for-assessing-fit)

## Appendix

#####  Hatching date

```{r, message=FALSE, warning=FALSE, eval=FALSE}
Hatch.index.values = 
  seq(from =  min(df$Hatch), to = max(df$Hatch), length.out = 100)

## Extract get 'par.index' of interest
index.meta <- data.frame(
  Sex = rep(c("Female", "Male"), each = 4),
  ageclass = rep(c("CY1", "CY2", "CY3", "CY4plus"), 2)) %>% 
  ## add par.index
  left_join(., ddl$S[,c("par.index", "Sex", "ageclass")],
            by = c("Sex", "ageclass")) %>% 
  ## code additional vars
  mutate(dummy = paste0(Sex, ageclass)) 

## simplify, one par.index for each dummy var
model.indices <- sapply(unique(index.meta$dummy), function(x) {
  ## return first row matching the query
  index.meta[index.meta$dummy == x,"par.index"][1]
}) %>% 
  unlist()

## subset index.meta
index.meta <- index.meta[index.meta$par.index %in% model.indices,]

## get estimates from the model
pred.model <- covariate.predictions(
  supported.models, 
  data = data.frame(
    Hatch = Hatch.index.values, Hacth.index.sq = Hatch.index.values^2), 
  indices = model.indices %>% as.numeric(),
  drop = T)[["estimates"]]

## extend df with grouping variables
pred.model <- left_join(pred.model,
                        index.meta[,c("par.index", "ageclass", "Sex")],
                        by = "par.index") %>% 
  mutate(Group = paste0(Sex, ageclass))
head(pred.model)

p.hatch <- 
  ggplot(pred.model, aes(x = Hatch, y = estimate, group = Group)) +
  geom_line(size = 1.5,aes(colour = Sex)) +
  geom_ribbon(aes(ymin = lcl, ymax = ucl, colour = Sex), alpha = 0.2) +
  scale_colour_brewer(palette = "Set1") + 
  theme_classic() +
  theme(
    panel.border = element_rect(fill = NA),
    panel.background = element_rect(fill = "Grey95"),
    text = element_text(size = 12),
    legend.position = 'none',
    aspect.ratio = 1) +
  xlab("Standardised Hatching date") +
  ylab("Survival probability [mean \U00B1 CI]") +
  scale_y_continuous(expand = c(0,0), 
                     limits = c(0,1),
                     breaks = seq(0, 1, .2)) +
  facet_wrap(~ageclass)
p.hatch
```

##### Morph

```{r, message=FALSE, warning=FALSE, eval=FALSE}
Morph = 2:4

## Extract get par.index of interest
index.meta <- data.frame(
  ageclass = rep(c("CY1", "CY2", "CY3", "CY4plus"), 1)) %>%   
  ## add par.index
  left_join(., ddl$S[,c("par.index", "ageclass")]) %>% 
  ## code additional vars
  mutate(dummy = paste0(ageclass)) 

## simplify, one par.index for each dummy var
model.indices <- sapply(unique(index.meta$dummy), function(x) {
  ## return first row matching the query
  index.meta[index.meta$dummy == x,"par.index"][1]
}) %>% 
  unlist()

index.meta <- index.meta[index.meta$par.index %in% model.indices,]

pred.Morph <- covariate.predictions(
  supported.models, 
  data = data.frame(
    Morph = Morph, bl.index.sq = Morph^2), 
  indices = index.meta$par.index)[["estimates"]]

## extend df with grouping variables
pred.model <- left_join(pred.Morph,
                        index.meta[,c("par.index", "ageclass")],
                        by = "par.index") %>% 
  mutate(Group = ageclass)

pred.model <- left_join(index.meta, pred.model[,c("par.index", "Morph", "estimate", "se", "lcl", "ucl", "ageclass", "Group")]) %>% 
  unique.data.frame()

pred.model$Morph <- factor(pred.model$Morph,
                           labels = c("Dark", "Intermediate", "Light") )

p.Morph <- 
  ggplot(pred.model, aes(x = as.factor(Morph),
                         y = estimate,
                         group = ageclass)) +
  geom_errorbar(aes(ymin = estimate - se, ymax = estimate + se), width = .25, size = 1.5) +
  geom_point(size = 3, colour = "Black", shape = 15) +
  scale_colour_brewer(palette = "Set1") + 
  theme_classic() +
  theme(
    panel.border = element_rect(fill = NA),
    panel.background = element_rect(fill = "Grey95"),
    text = element_text(size = 12),
    legend.position = 'bottom',
    aspect.ratio = 1) +
  xlab("Plumage Morph") +
  ylab("Survival probability [mean \U00B1 SE]") +
  facet_wrap(~ageclass) +
  scale_y_continuous(expand = c(0,0), 
                     limits = c(0,1),
                     breaks = seq(0, 1, .2)) 

p.Morph
```


##### Brood Rank

```{r, message=FALSE, warning=FALSE}
Rank = 1:4

## Extract get par.index of interest
index.meta <- data.frame(
  ageclass = rep(c("CY1", "CY2", "CY3", "CY4plus"), 1)) %>%   
  ## add par.index
  left_join(., ddl$S[,c("par.index", "ageclass")]) %>% 
  ## code additional vars
  mutate(dummy = paste0(ageclass)) 

## simplify, one par.index for each dummy var
model.indices <- sapply(unique(index.meta$dummy), function(x) {
  ## return first row matching the query
  index.meta[index.meta$dummy == x,"par.index"][1]
}) %>% 
  unlist()

index.meta <- index.meta[index.meta$par.index %in% model.indices,]

pred.rank <- covariate.predictions(
  supported.models, 
  data = data.frame(
    Rank = Rank, bl.index.sq = Rank^2), 
  indices = index.meta$par.index)[["estimates"]]

## extend df with grouping variables
pred.model <- left_join(pred.rank, index.meta[,c("par.index", "ageclass")], by = "par.index") %>% 
  mutate(Group = ageclass)

pred.model <- left_join(index.meta, pred.model[,c("par.index", "Rank", "estimate", "lcl", "ucl", "ageclass", "Group")]) %>% 
  unique.data.frame()

p.rank <- 
  ggplot(pred.model, aes(x = Rank, y = estimate, group = ageclass)) +
  geom_errorbar(aes(ymin = lcl, ymax = ucl), width = .25, size = 1.5) +
  geom_point(size = 3, colour = "Black", shape = 15) +
  scale_colour_brewer(palette = "Set1") + 
  theme_classic() +
  theme(
    panel.border = element_rect(fill = NA),
    panel.background = element_rect(fill = "Grey95"),
    text = element_text(size = 12),
    legend.position = 'bottom',
    aspect.ratio = 1) +
  xlab("Brood Rank") +
  ylab("Survival probability [mean \U00B1 CI]") +
  facet_wrap(~ageclass)

p.rank
```


## Literature cited

[^1]: Lebreton, J. D., Burnham, K. P., Clobert, J., & Anderson, D. R. (1992). Modeling survival and testing biological hypotheses using marked animals: a unified approach with case studies. Ecological monographs, 62(1), 67-118.
[^2]: White, G. C., & Burnham, K. P. (1999). Program MARK: survival estimation from populations of marked animals. Bird study, 46(sup1), S120-S139.
[^3]: Laake, J. L. (2013). RMark: an R interface for analysis of capture-recapture data with MARK.
[^4]: Burnham, K., & Anderson, D. (2002). Model selection and multi‐model inference: A practical information‐theoretic approach (2nd ed.). New York, NY: Springer.
[^5]: McClintock, B. T., & White, G. C. (2012). From NOREMARK to MARK: software for estimating demographic parameters using mark–resight methodology. Journal of Ornithology, 152(2), 641-650.
[^6]: Afroz, F., Parry, M., & Fletcher, D. (2019). Estimating overdispersion in sparse multinomial data. Biometrics.
[^7]: Reneerkens, J., Versluijs, T. S., Piersma, T., Alves, J. A., Boorman, M., Corse, C., ... & Ntiamoa‐Baidu, Y. (2019). Low fitness at low latitudes: Wintering in the tropics increases migratory delays and mortality rates in an Arctic breeding shorebird. Journal of Animal Ecology.
[^8]:Hurrell, J.W., Kushnir, Y., Ottersen, G. & Visbeck, M. (2003). An overview of the North Atlantic oscillation. The North Atlantic Oscillation: climatic significance and environmental impact, 134, 1–35.